### 2016-10-14 documentation/paper.tex
More discussion in doc

\documentclass[a4paper,11pt,twoside]{article}

\makeatletter

\newcommand\frontmatter{%
    \cleardoublepage
  %\@mainmatterfalse
  \pagenumbering{roman}}

\newcommand\mainmatter{%
    \cleardoublepage
 % \@mainmattertrue
  \pagenumbering{arabic}}

\newcommand\backmatter{%
  \if@openright
    \cleardoublepage
  \else
    \clearpage
  \fi
 % \@mainmatterfalse
   }

\makeatother

\usepackage[utf8]{inputenc}
\usepackage{verbatim}
\usepackage[british]{babel}

\usepackage{url}

\usepackage{amsmath}
\usepackage{amsthm}

\usepackage[mathlf,fourierbb,minionint,opticals,fullfamily,swash]{MinionPro}
%\usepackage[mathlf,fourierbb,minionint]{MinionPro}
\usepackage{MnSymbol}
\usepackage{pifont}
%\usepackage[all]{xy}
\usepackage{graphicx}
%\usepackage[dvips]{graphicx}

\newcommand{\name}[1]{\textsf{#1}}

\usepackage[activate]{pdfcprot}
\usepackage[active]{srcltx}
\usepackage{fancyhdr}
\pagestyle{fancy}
%\fancyhead[RO,LE]{\thepage}
\fancyhead{}
\fancyhead[LE]{\nouppercase{\textssc{\leftmark}}}
\fancyhead[RO]{\nouppercase{\textssc{\rightmark}}}
\fancyfoot[C]{\thepage}
%\fancyfoot[C]{\Pisymbol{MinionPro-Extra}{120} \thepage\ \Pisymbol{MinionPro-Extra}{120}}
%\fancyfoot[LO]{\thepage}
%\fancyfoot[RE]{\thepage}
%\fancyfoot{}

\usepackage{enumitem}

\usepackage{listings}
\lstdefinestyle{customc}{
  belowcaptionskip=1\baselineskip,
  escapeinside=``,
  breaklines=true,
  xleftmargin=\parindent,
  showstringspaces=false,
  basicstyle=\small\ttfamily,
  columns=fullflexible,
  keepspaces=true,
}

\newcommand{\eq}{\ensuremath{\sim}}

\usepackage[explicit,toctitles]{titlesec}

\usepackage[table]{xcolor}

\hyphenation{Doth-ra-ki}

\title{The Warrior Stabs the Goat: Creating a Dothraki Resource Grammar with Grammatical Framework}
\author{Benno Kuckuck}

\begin{document}
\frontmatter
\begin{titlepage}
\clearpage{\pagestyle{empty}\cleardoublepage}
\phantom{bla}
%\vfill
\begin{center}
\includegraphics[width=5cm]{hhu-logo.jpg}

\vspace{2cm}

{\Large Term Paper}
\vspace{1cm}
\hrule
\vspace{0.4cm}
\textsc{\Huge The Warrior Stabs the Goat}
\vspace{0.5cm}

{\LARGE Creating a Dothraki Resource Grammar with Grammatical Framework}

\vspace{0.4cm}
\hrule
\vspace{2.5cm}

{\Large Maria Henkel}\\
2008709 $\bullet$ maria.henkel@hhu.de
\vspace{0.5cm}

{\Large Benno Kuckuck}\\
2340409 $\bullet$ benno.kuckuck@hhu.de
\vspace{1.5cm}

{\Large 17 October 2016}

\vspace{2.5cm}

\vfill
Lecturers:  Yulia Zinova, Rainer Osswald\\
Course: Introduction to ``Grammatical Framework'', SS 2016

\end{center}
\vfill
\end{titlepage}

\tableofcontents

\newpage

\thispagestyle{empty}
\begin{center}
\Large
  \includegraphics[width=5cm]{drogo.png}
  \vspace{1.5cm}

  \begin{minipage}{10cm}
  \centering
  ``Khal Drogo shillo chiories she krazaaj. Chiori fin vijazer 
  adraes zoqwe Khales Drogo. Khal Drogo vinde ma chiories ma 
  adraes she krazaaj. Khal Drogo vachrara.''

  \vspace{0.7cm}
  \hrule
  \vspace{0.3cm}

  \color{gray}
``Khal Drogo met a woman on a mountain. The woman, who 
protected a turtle, kissed Khal Drogo. Khal Drogo stabbed the 
woman and the turtle on the mountain. Khal Drogo stinks.''

\vspace{0.5cm}
\color{black}
GF Run-Time System, 2016
  \end{minipage}
\end{center}

\newpage

\thispagestyle{empty}

\mainmatter
\section{Introduction}
Grammatical Framework (GF) is a functional programming language, 
created in 1998 \cite{gf-book}, doubling as a categorial grammar 
formalism that can be used to implement (natural) language grammars. 
As GF is ``working from a language-independent representation of meaning''
\cite{wiki-gf}
it can be used to generate or parse text in different languages, 
e.g.\ for use in machine translation. 
Since GF uses a symbolic approach to processing and translating language,
in most cases its translations are more precise than those of the 
more commonly used statistical tools \cite{brown-frederking}. 
For example, on somewhat complicated input,
the ``Google Translate'' service \cite{google} will in many cases
produce ungrammatical sentences, while the GF translator 
\cite{gf-translate} has no problems producing 
a correct translation:

\begin{center}
  \rowcolors{2}{gray!25}{white}
  \begin{tabular}{p{3.7cm}|p{4cm}|p{3.7cm}}
  \rowcolor{gray!50}
    Input & Output Google Translate & Output GF Translator\\\hline
    I heal these women. & 
    Ich diese Frauen zu heilen. & 
    Ich verheile diese Frauen.\\
    The warrior will kiss that cheese behind the tree. & 
    Der Krieger wird Kuss, der K\"ase hinter dem Baum. & 
    Der Krieger wird jenen K\"ase hinter dem Baum k\"ussen.\\
    Diese hei\ss e Frau heilt diese K\"onigin im Meer. & 
    This hot woman heals the queen of the sea.  & 
    This hot woman heals this queen in the sea.
  \end{tabular}
\end{center}

Symbolic formalisms like GF -- unlike Google Translate -- depend 
on a large knowledge base of predefined grammar rules 
to achieve high-quality language processing: Its 
own standard library, the GF Resource Grammar Library, covers 
more than 38 different languages \cite{gf-status} -- although many of 
them are not yet complete -- and is still growing. 
In 2010, it already consisted of more than 500,000 
lines of code \cite{gf-book}. 
As Grammatical Framework is an open-source project \cite{ranta-rgl} 
and still under development, everyone is invited to use it 
or even help further develop it, for example by adding languages to 
the library, which is the aim of the project described in this paper. 
Our goal is to write a GF resource grammar for the Dothraki 
language and, eventually, be able to add it to the GF 
Resource Grammar Library. The work described in this paper is meant to 
lay a foundation for this goal (see Section \ref{sec-goals}).
Dothraki is a constructed language, developed by the language creator and writer 
David J. Peterson \cite{peterson-book}. The Dothraki 
are a fictional race of nomadic horse warriors in the ``A Song of Ice and 
Fire'' book series by George R. R. Martin \cite{martin}. 
As the series was adapted for television by the television network HBO, 
the language bits found in Martin's books had to be extended 
to a fully functional language \cite{peterson-web}: The Dothraki language. 
Today, it has a vocabulary of over 1,400 words 
\cite{vocab}. Examples of 
Dothraki language grammar will appear
in Section \ref{sec-impl}, but much more information can 
be found in the book ``Living Language Dothraki'' \cite{living-language}, 
on Peterson's blog \cite{peterson-web} and on the Tongues of Ice and 
Fire Wiki (2016) among many other sources.

This paper aims to serve as a short documentation and explanation of our 
project. It will cover the project aim and scope, as well as its 
structure and an exposition of our implementation of
Dothraki grammar rules in GF (Section \ref{sec-impl}). 
We will also discuss problems and limitations and briefly describe 
future plans before we summarize and evaluate the project in its 
current state and our experience while working on it (Section \ref{sec-discuss}).

\section{The Project}\label{sec-project}
As mentioned before, the eventual aim of this project is to create a 
resource grammar for the Dothraki language and add it 
to the GF Resource Grammar Library, so that other Dothraki fans 
and/or GF enthusiasts may use it. While there are currently a 
few Dothraki language learning apps and online translators 
available (see \cite{cognitus}, \cite{funtrans}), 
these are very limited, primarily relying on word-for-word 
translation unaware of any grammar rules. An open-source 
Dothraki resource grammar for GF, which already provides a wide 
range of tools and options for integrating with other projects, 
would undoubtedly be a significant addition to the available range 
of Dothraki language resources.

\subsection{About Resource Grammars and the Standard Library}\label{sec-rgl}

GF, like other symbolic approaches to machine translation and language 
processing, is well-suited for providing high-quality output in more 
narrowly focused applications (as opposed to statistical approaches, whose 
strengths lie more in providing acceptable output in a very wide range 
of scenarios). Most applications however share a large part of the rules 
necessary to parse or construct correct sentences, namely those that are 
concerned with the morphology and syntax of the involved language(s). 
While these could be implemented ad-hoc for every application, it is much 
more economic to provide language-specific but application-independent 
rules as a reusable library. This is the goal of GF's Resource Grammar 
Library.

The idea is to free application programmers from having to deal with 
``specialized linguistic expertise'' (Ranta, 2009). 
By providing a common interface across languages, it becomes 
considerably easier (in some cases even trivial) for the 
application programmer to add support for additional languages. 
Considerable work has been and is being done by the GF developers 
and resource grammar programmers to make this possible:

\begin{quote}
The idea of using a grammar as a software library is, to our knowledge,
new in GF. It has required a considerable effort in the design 
and implementation of the GF programming language: a type system, 
a module system, compilation techniques, and optimizations. [...] 
The  effort
made in this work is supplemented by a substantial effort in the 
library itself.
The code included in the library is more than twice in size compared to the
implementation of GF.\hfill\cite{ranta-rgl}
\end{quote} 
This division of labour benefits not just the application programmers,
however, but also the linguistics experts implementing the 
application-independent libraries:\\ Firstly, because a resource 
grammar, once written, can ``serve an unlimited number of applications'' 
\cite[p.\ 200]{gf-book}, but also because ``the existing library 
specification will help to identify the linguistic issues 
and avoid pitfalls'' by providing a well thought-out, 
language-independent interface.

\subsection{Project Goals and Structure}\label{sec-goals}
While in the long-term, our work is aimed towards a full 
implementation of a Dothraki resource grammar, this was not a 
realistic goal for this term paper, for two main reasons: 
First, Ranta estimates that an interval of three to six months 
of full-time work is needed for an experienced GF programmer to 
implement a new resource grammar \cite[p.\ 222]{gf-book}, which far 
exceeds the time frame for this term paper. And secondly, 
the Dothraki language itself is still a work-in-progress, with 
significant parts of the grammar either as yet undocumented, or 
possibly non-existent, which at present makes a full implementation 
of all the grammatical functions covered by GF's RGL arguably 
impossible (for more on this point, see the discussion in Section \ref{sec-discuss}).

In Ranta's book \cite[pp.\ 237--245]{gf-book} (see Figure \ref{fig-ita}), a mini resource 
grammar for Italian is implemented as an example.

\begin{figure}[h]
\centering
  \includegraphics[width=10cm]{ita.png}
  \caption{\label{fig-ita}Mini Resource Grammar Modules and Dependencies \cite[p.\ 209]{gf-book}}
\end{figure}

Our work lies somewhere between this example and a full resource grammar. 
Like in the Italian example, the abstract grammar specification 
our resource grammar is built on, is only a subset of that of the full 
RGL. Unlike the Italian example, it adheres much more closely to the
structure of the full RGL (see Figure \ref{fig-rgl}), employing the same file structure 
and API, in hopes of easing a later expansion to a full resource grammar.
It is also much more extensive than Ranta's example, covering a much wider
range of grammatical constructions.\footnote{As a quick comparison, the example 
from the book implements 12 grammatical categories, whereas our Dothraki 
implementation covers 27 of the 48 categories of the full RGL. In terms 
of API methods implemented, the comparison is even starker: Ranta's 
example grammar implements 19 API methods, our grammar 
implements almost 200.}

\begin{figure}[h]
\centering
  \includegraphics[width=13.5cm]{rgl.jpg}
  \caption{\label{fig-rgl}Module Structure of a Full Resource Grammar \cite{gf-slides}}
\end{figure}

Just like the GF repositories \cite{detrez-camilleri}, 
our full project is hosted on GitHub \cite{github} so that others may see, use and contribute to it.

The GF Resource Grammar Library consists essentially of three parts:
\begin{enumerate}
  \item The \emph{API} is the outward-facing part of the library.
  It contains those modules which will be used by application
  grammar programmers. It, in turn, is composed of two parts:
  \begin{itemize}
    \item A (large) language-independent
  part (contained in the \name{api} folder of the RGL), primarily
  the module \name{Syntax}, which (roughly) consist of the module
  \name{Cat}, defining the abstract categories
  of GF's langauge-independent representation of syntax
  and the module \name{Constructors}, which defines
  a wide array of functions for building up larger syntactical
  components (e.g.\ noun phrases, verb phrases, clauses) 
  from constituent parts.
  \item A (smaller) part specific to each language of the RGL, 
  defined in a module called \name{Paradigms\emph{XXX}},
  which contains functions for constructing instances of the
  basic lexical categories (e.g.\ nouns, verbs, adjectives).
  \end{itemize}

  \item The API functions are defined in terms of more basic
  operations defined in the \emph{abstract part} of the RGL,
  \footnote{For example the API contains 22 overloads for the function
  \name{mkCl}, used to construct clauses, but almost
  all of these are based on the single method 
  \name{predVP} from module \name{Sentence}.}
  which is contained in the folder \name{abstract}

  These are the operations which must be implemented for each 
  language.

  \item The \emph{concrete part} of the RGL, consists
  of concrete implementations of all abstract modules
  for each language that is part of the RGL. These are mostly contained
  in the folders carrying the name of the language.

  Some of the abstract operations and categories are implemented similarly
  for almost all languages. Such implementations are contained
  in the \name{common} folder.

  \item Additionally, but optionally, many languages
  also supply a large dictionary of words available in
  that language (allowing application grammar writers
  to use e.g.\ \verb+rise_N+ directly, instead of having to use
  \verb+mk_V "rise" "rose" "risen"+ and risking getting
  the irregular forms wrong).
\end{enumerate}

As mentioned before, our project structure adheres closely
to this template (see Figure \ref{fig-folders}):

\begin{itemize}
  \item The \name{simpleapi} and \name{simpleabstract} folder contain
  the API part and abstract part respectively.

  Their contents are selectively copied over from the corresponding
  files in the \name{api} and \name{abstract} folders of the full RGL.

  This allows us to build a working resource grammar, without
  having to implement every single category and operation
  first. While our Dothraki resource grammar grew and continues
  to grow these are gradually expanded, until they converge
  towards the full RGL.

  \item The \name{simpleenglish} folder conatins those
  parts of the English resource grammar, which are relevant
  for our trimmed-down API.

  \item The \name{common} folder is taken straight from
  the full RGL.

  \item Finally, the \name{simpledothraki} folder contains
  our original contribution, the (somewhat
  trimmed-down) resource grammar for Dothraki.
\end{itemize}

There are three more folders, which are not part of the resource grammar:
\begin{itemize}
  \item The folder \name{example} contains an application grammar
  that uses our Dothraki resource grammar and the English resource
  grammar.

  \item The folder \name{documentation} contains this documentation.

  \item The folder \name{vocabulary} contains a Python script
  used for scraping the wiki source code of 
  the publically available vocabulary list \cite{vocab} and
  automatically generating a GF-compatible Dothraki dictionary
  (see Section \ref{sec-vocab}).
\end{itemize}

\begin{figure}[h]
\begin{itemize}[itemsep=0.5pt]
\item simpleapi
\item simpleabstract
\item simpledothraki
\item simpleenglish
\item common

\vspace{0.2cm}
\item example

\vspace{0.2cm}
\item documentation 
\item vocabulary
\end{itemize}
\caption{\label{fig-folders}The project's folder structure}
\end{figure}

\subsection{Implementation of Grammatical Categories in Dothraki}
\label{sec-impl}
In this section we will go into some of the design decisions 
involved in the implementation of the Dothraki resource grammar. 
In order to get a high-level overview, we will first take a look 
at the linearizations of some of the categories defined in the 
API part of the abstract resource grammar library, showing how 
Dothraki grammar can be fit into the GF scheme.

\subsubsection{Nouns and Noun Phrases}
The relevant categories here are \name{N} (nouns), 
\name{CN} (common nouns), \name{NP} (noun phrases), \name{Pron} (pronouns) 
and \name{PN} (proper names), which are realized as 
follows:\footnote{Here, and in the following examples,
the code snippets do not necessarily appear exactly like
this in the actual source code, since usually some parts are 
defined in the \name{ResDot} 
module and some in \name{CatSimpleDot}, for reasons of 
reusability and implementation hiding. However, in all cases,
the actual implementations are equivalent to the listings
supplied here.}

\lstset{style=customc}
\lstMakeShortInline{§}

\begin{lstlisting}
N, CN = {s : Number => Case => Str ; a : Animacy } ;
NP, Pron = {s : Case => Str ; agr : Agr } ; 
PN = {s : Case => Str } ;
\end{lstlisting}

These implementations follow in a fairly straightforward fashion 
from the basic features of nouns in the Dothraki language: 
In Dothraki, every noun falls into one of two classes, 
\emph{animate} and \emph{inanimate} (which play a very similar 
role to grammatical gender in other languages). 
The language uses five cases: nominative, accusative, 
genitive, ablative and allative. Hence, we define:

\begin{lstlisting}
Animacy = Anim | Inanim ;
Case = Nom | Gen | Acc | All | Abl ;
\end{lstlisting}

Noun phrases, as well as pronouns, have a fixed number and person, 
which is stored in the §agr§ field of type §Agr§, simply defined as follows:

\begin{lstlisting}
Agr = Ag Person Number ;
\end{lstlisting}

There does not currently seem to be a need to store animacy 
in the agreement record. This is unlike English, where the 
§NP§s gender can affect the verb phrase 
(``The boy loves himself.'' but ``The girl loves herself.'') 
and thus needs to be stored in the §NP§. 
Finally, proper names are (so far as we can tell) 
always animate and third person singular, so none of that 
information needs to be stored in §PN§.

In order to turn a common noun into a noun phrase, 
we usually need a determiner. The relevant categories here are 
§Num§ (numerals), §Quant§ (quantifiers), §Predet§ (predeterminers) 
and §Det§ (determiners, roughly composed of optional predeterminers, 
a quantifier and a numeral). 
Numerals in the proper sense, as well as predeterminers, 
are not yet implemented in our resource grammar 
(there is a lack of good information on determiner structure in 
Dothraki), but since the §Num§ category is used in the 
GF Resource Grammar Library not just for representing 
numerals but also to determine the number of 
the determiner, we have a dummy implementation in 
our resource grammar:

\begin{lstlisting}
Num = {s: Str ; n : Number} ;
Quant = {s : QuForm => Case => Str } ;
Det = {s : Animacy => Case => Str ; n : Number ; s2 : Str } ;
\end{lstlisting}

The §s§ field of §Num§ is always empty in our implementation, 
this is necessary for technical reasons related to the implementation 
of GFs parser \cite{google-groups}.

Quantifiers in Dothraki (such as ``jin'' \eq\ ``this'' and ``rek'' \eq\ ``that'') are inflected for animacy, number and case, but the singular and plural inanimate forms are always identical, so ‘QuForm’ is defined as follows:

\begin{lstlisting}
QuForm = QAnim Number | QInanim ;
\end{lstlisting}

Unlike a §Quant§, a §Det§ has a fixed number.
An interesting feature of determiners in Dothraki is 
that they can appear either before the noun 
(``jinak adra'' = ``this turtle'') or after (such as determiners built 
from possessive pronouns: ``adra anni'' = ``my turtle''). 
Conceivably,\footnote{we cannot currently say for certain, 
due to lack of examples and documentation on 
Dothraki determiner structure} a determiner could even be split 
into two parts (``my two turtles'' $\stackrel{?}=$ ``akat adrasi anni''). 
In anticipation of this, §Det§ has two §Str§ components, §s§ and §s2§. 
Since possessive pronouns are the only examples of postposed 
determiners we know so far, and these happen not to 
inflect with animacy or case, the §s2§ part is simply a §Str§, whereas 
the §s§ part is a table of inflected forms.

\subsubsection{Verbs and Verb Phrases}\label{sec-verbs}

The relevant categories here are §V§ (intransitive verbs), 
§V2§ (transitive verbs with an §NP§ complement) 
and §VP§ (verb phrases), whose lincats are as follows:

\begin{lstlisting}
param VFormPN = Pers1 Number | Pers2 | Pers3 Number ;
param VForm = 
APast Polarity Number 
  | APresent Polarity VFormPN 
  | AFuture Polarity VFormPN 
  | ImpFormal Polarity 
  | ImpInformal Polarity ;

oper Verb : Type = {
  s : VForm => Str ; 
  inf : Str ; 
  part : Str
  } ;

V = Verb ;
V2 = Verb ** {objCase : Case} ;
VPSlash = Verb ** {objCase : Case ; subjpost : Str} ;
VP = Verb ** {compl : Str; subjpost : Str} ;
\end{lstlisting}

The basic structure §Verb§ contains the infinitive and participle 
as well as all conjugated forms of a verb. 
The somewhat convoluted setup of the §VForm§ parameter type is 
owed to the peculiarities of conjugation in Dothraki: 
First off, somewhat unusually,\footnote{from the Indo-European
viewpoint, that is, not for natural languages in general} in all 
tenses and moods, 
verb forms vary with the polarity of the sentence 
(``Me dothra\textbf{e}'' = ``He rides'', ``Me vos dothra\textbf{o}'' = ``He does not ride''). 
In the present and future tenses, verbs are inflected for 
person and number, but the second person singular and plural 
forms are always identical (hence the definition of §VFormPN§). In 
the past tense, verbs are inflected for 
number only, not for person (but unlike in the present and 
future tenses, second person singular and plural forms do differ). 
There are also formal and informal imperative forms.

Transitive verbs are basically the same as intransitive verbs, 
but they also assign a case to their §NP§ complement (mostly 
accusative, but other cases do appear), which is 
stored in the §objCase§ field.

A verb phrase essentially consists of a verb with a 
complement. There is also a field §subjpost§ which again encodes 
a peculiar feature of Dothraki grammar: 
A number of constructions, which, in other languages, 
are rendered by verbal auxiliaries (can, must) or verbs with §VP§ complements 
(try to), are realized in Dothraki 
via non-inflecting particles (e.g.\ ``laz'' \eq\ ``can'' 
and ``eth'' \eq\ ``must''), which do not otherwise affect 
the conjugation of the main verb (``Me dothrak'' = ``He rides'', 
``Me laz dothrak'' = ``He can ride'', 
``Me kis dothrak'' = ``He tries to ride'')\footnote{This could be compared
to the way negation is expressed in many Indo-European languages,
namely by a non-inflecting particle (``not'', ``nicht'', ``ne \dots\ pas'', etc.),
which is adposed to the verb in some manner, but does not otherwise
affect inflection.}. These have to 
be regarded as part of the verb phrase but, 
syntactically, are not preposed to the verb phrase but postposed 
to the subject of the clause (this makes a difference, e.g., in 
relative clauses, which have VSO word order instead of the 
standard SVO word order). The same is true of passive constructions, 
which are indicated by ``nem'' postposed to the subject 
(``Lajak dothrae'' = ``The warrior rides'', `` Hrazef nem dothrae'' = 
``The horse is ridden''; note that the verb still agrees in number and person
with the syntactical subject of the clause).

§VPSlash§ represents a verb phrase missing a complement 
and structurally kind of sits between §V2§ and §VP§. 
It is generated from a §V2§ (via §SlashV2 : V2 -> SlashVP§) 
and can then either be turned into a §VP§ by adding an object 
(via §ComplSlash : VPSlash -> NP -> VP§’) or into a §ClSlash§ (clause 
missing an object) by adding a subject (via 
§SlashVP : VPSlash -> NP -> ClSlash§), to be used, e.g., in a relative 
sentence (``the goat which the warrior stabbed''). As such it does 
not have a complement, like §VP§, and retains the §objCase§ from §V2§, 
but it does have the §subjpost§ field from §VP§ (so we can form, e.g., 
``the goat which the warrior can stab'' = ``dorvi fin vindee lajak laz'').

In order to implement verbal auxiliaries such as ``can'' and ``must'', 
which are represented as §VV§s in GF, we also have a 
dummy lincat for §VV§:

\begin{lstlisting}
VV = {s : Str} ;
\end{lstlisting}

Since ``can'', ``must'' etc.\ are simply uninflected particles 
in Dothraki, this is enough to represent those. 
This should of course not be considered a final design, 
since it is unlikely that all verbs with verb phrase 
complement can be realized like this in Dothraki. 
However, there is currently very little documentation available 
on how other §VV§ verbs are realized, so this dummy 
implementation will have to do for now.

\subsubsection{(Relative) Clauses and Sentences}

There are no big surprises in the lincats of §Cl§ and §QCl§ 
(clauses and question clauses, respectively) or §S§ and §QS§ 
(sentences and question sentences):

\begin{lstlisting}
S, QS = {s : Str} ;
Cl, QCl = {s : Tense => Anteriority => Polarity => Str} ;
\end{lstlisting}

The lincats of these categories are fairly canonical in 
many languages and in Dothraki they are basically the same as 
in English or German. The main differences between these 
languages come from whether the linearization also depends 
on some kind of §Order§ parameter, for example distinguishing 
indirect from direct questions or subordinate from 
main clauses. It might be necessary to add something 
like that to our lincat later, if more information about, e.g.,
indirect questions in Dothraki becomes available, 
but for now there is no indication that we 
need such an extra parameter.

Relative clauses are more interesting. 
The relevant categories here are §RP§ (relative pronoun), §ClSlash§ 
(a clause missing an object), §RCl§ (relative clauses) and §RS§
(relative sentence):

\begin{lstlisting}
RP = {s : QuForm => Case => Str } ;
RCl = {s : Tense => Anteriority => Polarity => Animacy => Number => Str } ;
ClSlash = {s : Tense => Anteriority => Polarity => Str; 
    subj: Str ; objCase : Case} ;
RS = {s : Animacy => Number => Str } ;
\end{lstlisting}

Relative pronouns in Dothraki inflect according to the 
animacy and number of the noun phrase they attach to, 
and the case that the relativized noun phrase would have 
been assigned in the embedded relative sentence (similar to German). 
Hence, the linearization of §RS§ depends on the 
animacy and number of the noun phrase 
it is eventually attached to. §RCl§ is to §RS§ 
what §Cl§ is to §S§, i.e., it is a relative sentence, 
whose tense, anteriority and polarity have not yet 
been determined. §ClSlash§ has to remember the case of 
the missing object, so it can assign it to the relative 
pronoun, if used in a relative sentence 
(via §RelSlash : RP -> ClSlash -> RCl§).

There is one more important difference between §Cl§ and §ClSlash§. 
When a clause is constructed from a verb phrase and 
subject via §PredVP : NP -> VP -> Cl§, the noun phrase is 
just prepended to the verb phrase. We cannot do this 
when constructing a §ClSlash§ via §SlashVP : NP -> VPSlash -> ClSlash§, 
because of a quirk in Dothraki word order: 
Whereas regular main clauses and questions 
follow SVO word order, relative clauses have 
VSO word order. So depending on whether a §ClSlash§ is eventually 
turned into a relative sentence (via §RelSlash : RP -> ClSlash -> RCl§) 
or a question (via §QuestSlash : IP -> ClSlash -> QCl§), we have 
to choose the word order accordingly 
(``dorvi fin vindee lajak'' = ``the goat which the warrior stabs'' vs 
``Fin lajak vindee?'' = ``What does the warrior stab?''). 
Therefore, §ClSlash§ contains the (linearized) subject in 
a field, §subj§, instead of embedding it in the §s§.

\subsubsection{Adjectives}
The relevant categories here are §A§ (adjectives), §AP§ 
(adjectival phrases) and §Comp§ (''complement of a copula'', 
though see below):

\begin{lstlisting}
A = {s : Degree => Number => ACase => Str ; 
    pred : VForm => Str };
AP = {s : Number => ACase => Str ; pred : VForm => Str } ;
Comp = {s : VForm => Str} ; 
\end{lstlisting}

There are two basic uses for adjectival phrases: 
attributive (as in ``the strong warrior'') and predicative 
(as in ``the warrior is strong''). Attributive use works 
similarly in Dothraki to, e.g., English and German: The 
adjectival phrase inflects for number and case in agreement 
with the noun it modifies. In Dothraki the forms for all 
the cases other than nominative are always identical, 
however, so we define

\begin{lstlisting}
ACase = ANom | AOther ;
\end{lstlisting}

instead of using §Case§ and having to duplicate these forms 
in the table.

In English or German, predicative use of an adjective 
employs a copula (``to be'' or ``sein''), which is inflected for 
tense, whereas the adjective is invariant (the same happens in 
sentences like ``the woman is a warrior'' or ``the man is on 
the mountain'', where the copula ``to be''
connects two noun phrases or a noun phrase and an adverbial phrase). 
Dothraki, however, is an 
entirely copulaless language. When an adjective is used 
predicatively, the adjective itself is essentially turned 
into a verb and inflected for tense and polarity:
\begin{quote}
``the strong warrior'' =
``lajak haj''\\
``The warrior is strong'' = ``lajak haja''\\
``the warrior will be strong'' = ``lajak vahaja''\\
``the warrior was not strong'' = ``lajak vos ahajo''
\end{quote}

Similarly, in a sentence like 
``the woman is a warrior'', the §NP§ itself inflects for tense,
with the ablative and allative form indicating past 
and future tense, respectively: 
\begin{quote}
``The woman is a warrior'' = ``Chiori lajak''\\
``The woman was a warrior'' = ``Chiori lajakoon''\\
``The woman will be a warrior'' = ``Chiori lajakaan''
\end{quote}

As a result, the category §Comp§ (i.e., ``strong'' in ``the warrior 
is strong'' and ``a warrior'' in ``the woman is a warrior'') essentially 
behaves like an intransitive verb in Dothraki. 
§A§ and §AP§ have to contain all conjugated forms 
of the verb as it might be used in a predicative context.

Indeed our lincats for §A§, §AP§ and §Comp§, while very much unlike 
the lincats in the English and German resource grammars, 
are quite similar to the lincats in the resource grammar 
for Japanese, another language in which the adjective 
itself can be inflected for tense and polarity in predicative 
use via a complex set of suffixes. As in our resource 
grammar, this leads to §AP§ and §Comp§ being structurally 
very similar to §VP§:
\begin{lstlisting}
VP = {
    verb : Speaker => Animateness => Style => TTense => Polarity => Str ;  
    a_stem, i_stem : Speaker => Animateness => Style => Str ;  
    te, ba : Speaker => Animateness => Style => Polarity => Str ;
    prep : Str ; obj : Style => Str ; 
    prepositive : Style => Str ; needSubject : Bool} ;
Comp = {
    verb : Animateness => Style => TTense => Polarity => Str ;
    a_stem, i_stem : Animateness => Style => Str ; 
    te, ba : Animateness => Style => Polarity => Str ;
    obj : Style => Str ; prepositive : Style => Str ; 
    needSubject : Bool} ;
AP = {
    pred : Style => TTense => Polarity => Str ; 
    attr, adv, dropNaEnging, prepositive : Style => Str ; 
    te, ba : Style => Polarity => Str ; 
    needSubject : Bool} ;
\end{lstlisting}

\subsubsection{Other Categories}

Adverbial phrases, like in most other languages, are simply 
uninflected token strings, as implemented in §CommonX§:

\begin{lstlisting}
Adv = {s : Str} ; 
Prep = {s : Str ; c : Case} ;
\end{lstlisting}

Adverbial phrases can be formed from noun phrases 
using prepositions, which assign case in Dothraki, 
or sometimes are expressed \emph{only} by case. 
For example, §to_Prep§ from module §Structural§ is implemented 
in Dothraki simply as §{s = [] ; c = All}§: What is expressed in 
other laguages via the preposition ``to'', is understood in 
Dothraki from the modified noun phrase appearing in allative case.

Conjunctions in Dothraki have different forms 
depending on whether they are used to conjoin 
noun phrases/sentences, or as phrasal conjunctions to start a sentence:

\begin{lstlisting}
Conj = {s : Str; p : Str; n : Number} ;
\end{lstlisting}

E.g., the conjunction ``and'', when used to connect noun phrases,
is ``ma'' (as in ``ma lajak ma khaleesi'' = ``the warrior and the queen''), 
but translates as ``majin'', when used as a phrasal 
conjunction (''Majin lajak zoqwe khaleesies.'' = 
``And the warrior kissed the queen.'').

\begin{figure}
  \centering
  \includegraphics[width=13cm]{cats.png}
  \caption{\label{fig-cats}GF Resource Grammar Library Categories \cite{gf-slides}}
\end{figure}

Figure \ref{fig-cats} shows the main categories of the GF Resource 
Grammar Library and the current status of our implementation. 
The yellow categories have been fully or at least partially 
implemented in our resource grammar.

\subsection{Implementation of Syntactical Functions in Dothraki}

\subsubsection{Morphology}

Dothraki, generally speaking, has a much richer morphology than 
English, making the constructors for lexical categories 
supplied in §ParadigmsSimpleDot§ fairly complex. As an example, 
let us describe the construction of verbs 
(nouns and adjectives present similar challenges). 
Here is the constructor §mk2V : Str -> Str -> V§ 
which builds a §V§ from the infinitive form and the 
past singular (the two forms which would often be given in dictionaries):

\begin{lstlisting}
mk2V : Str -> Str -> Verb = \zalat,zal -> let {stem = stemV zalat zal} in {
  inf = zalat ;
  s = case stem of {
    fati@(fat + ("a"|"e"|"i"|"o")) => table {
      APast Pos Sg => zal ;
      APast Pos Pl => fati + "sh" ;
      APast Neg Sg => fat + "o" ;
      APast Neg Pl => fat + "osh" ;

      APresent pol pn => presForm fati pol pn ;

      AFuture Pos pn => presForm (pre {"a"|"e"|"i"|"o" => "v" ; _ => "a"} + fati) Pos pn ;
      AFuture Neg pn => presForm (pre {"a"|"e"|"i"|"o" => "v" ; _ => "o"} + fati) Neg pn ;

      ImpFormal Pos => stem ;
      ImpFormal Neg => fat + "o" ;

      ImpInformal Pos => fati + "s" ;
      ImpInformal Neg => fat + "os"
    } ;
    em => table {
      APast Pos Sg => zal ;
      APast Pos Pl => em + "ish" ;
      APast Neg Sg => em + "o" ;
      APast Neg Pl => em + "osh" ;

      APresent pol pn => presForm em pol pn ;

      AFuture Pos pn => presForm (pre {"a"|"e"|"i"|"o" => "v" ; _ => "a"} + em) Pos pn ;
      AFuture Neg pn => presForm (pre {"a"|"e"|"i"|"o" => "v" ; _ => "o"} + em) Neg pn ;

      ImpFormal Pos => em + "i" ;
      ImpFormal Neg => em + "o" ;

      ImpInformal Pos => em + "as" ;
      ImpInformal Neg => em + "os"
    } 
  } ;
  part = case stem of {
    fat + ("a"|"e"|"i"|"o") => stem + "y" ;
    _ => stem + "ay"
  }
} ;
\end{lstlisting}

Here §presForm : Str -> Polarity -> VFormPN -> Str§ is another 
function, that generates the present tense forms of a verb. 
This is reused for the future tense forms, which have 
the same endings, but are distinguished by a prefix.

\begin{lstlisting}
presForm : Str -> Polarity -> VFormPN -> Str = \stem,pol,pn -> case stem of {
  fati@(fat + ("a"|"e"|"i"|"o")) => case <pol,pn> of {
    <Pos, Pers1 Sg> => fati + "k" ;
    <Pos, Pers1 Pl> => fati + "ki" ;
    <Pos, Pers2> => fati + "e" ;
    <Pos, Pers3 Sg> => fati + "e" ;
    <Pos, Pers3 Pl> => fati + "e" ;

    <Neg, Pers1 Sg> => fat + "ok" ;
    <Neg, Pers1 Pl> => fat + "oki" ;
    <Neg, Pers2> => fati + "o" ;
    <Neg, Pers3 Sg> => fati + "o" ;
    <Neg, Pers3 Pl> => fati + "o"
  } ;
  em => case <pol,pn> of {
    <Pos, Pers1 Sg> => em + "ak" ;
    <Pos, Pers1 Pl> => em + "aki" ;
    <Pos, Pers2> => em + "i" ;
    <Pos, Pers3 Sg> => em + "a" ;
    <Pos, Pers3 Pl> => em + "i" ;

    <Neg, Pers1 Sg> => em + "ok" ;
    <Neg, Pers1 Pl> => em + "oki" ;
    <Neg, Pers2> => em + "i" ;
    <Neg, Pers3 Sg> => em + "o" ;
    <Neg, Pers3 Pl> => em + "i"
  }
} ;
\end{lstlisting}

As can be seen here, suffixes often differ according to 
whether the stem ends in a vowel or a consonant, which is 
handled by GF's pattern matching mechanisms. Future tense 
forms are indicated by a prefix (``v'' if the stem starts with a vowel, ``a''
if the stem starts with a consonant), which we implement employing 
GF's special ``prefix-dependent choice'' type.

In case the past singular form of the verb is not explicitly 
supplied, it needs to be guessed, which can be complicated:

\begin{lstlisting}
mk1V : Str -> Verb = \w -> case w of {
  ezo + "lat" => mk2V w ezo ;
  riss + "at" => mk2V w (addepenthesis riss) 
} ;
\end{lstlisting}

In general, the past singular form is just the stem of the verb, 
which mostly is derived by removing the "-lat" or "-at" 
infinitive ending.\footnote{Some verbs, such as "zalat" end in 
"-lat" but the "l" is part of the stem, 
not the infinitive ending. This cannot be guessed and has 
to be indicated in a dictionary.} However, this is complicated by the 
rules of epenthesis: The stem of "rissat" is "riss-", but the 
past singular is "risse", because a Dothraki word cannot end 
in "-ss". The rules for which consonant clusters may appear at 
the end of a word are complicated and for the most part captured in 
the function §addepenthesis : Str -> Str§, which adds an epenthesis 
"-e" to a string, if it ends in a disallowed consonant cluster:

\begin{lstlisting}
addepenthesis : Str -> Str = \w -> case w of {
  _ + ("a"|"e"|"i"|"o") => w ;
  _ + ("w"|"g"|"q") => w + "e" ;
  x + ("y"|"r"|"l") => case x of {
    _ + ("a"|"e"|"i"|"o") => w ;
    _ => w + "e"
  } ;
  x + ("m"|"n") => case x of {
    _ + ("a"|"e"|"i"|"o") => w ;
    y + ("w"|"y"|"r"|"l") => case y of {
      _ + ("a"|"e"|"i"|"o") => w ;
      _ => w + "e"
    } ;
    _ => w + "e"
  } ;
  x + ("th"|"s"|"sh"|"z"|"zh"|"kh"|"f"|"v") => case x of {
    _ + ("a"|"e"|"i"|"o") => w ;
    y + ("w"|"y"|"r"|"l"|"m"|"n") => case y of {
      _ + ("a"|"e"|"i"|"o") => w ;
      _ => w + "e"
    } ;
    _ => w + "e" 
  } ;
  x + ("j"|"ch"|"t"|"d"|"k") => case x of {
    _ + ("a"|"e"|"i"|"o") => w ;
    y + "w"|"y"|"r"|"l"|"m"|"n"|"th"|"s"|"sh"|"z"|"zh"|"kh"|"f"|"v") => case y of {
      _ + ("a"|"e"|"i"|"o") => w ;
      _ => w + "e"
    } ;
    _ => w + "e"
  }
} ;
\end{lstlisting}

This heuristic will give the right result in most (but not all) 
cases. Like in other resource grammars, there is always the 
option of supplying an irregular form by hand, if the heuristic fails.

\subsubsection{Syntax}

Many of the considerations about Dothraki syntax, 
that have gone into the design of our resource grammar, 
have already been mentioned in \ref{sec-impl}.
As one example of how the design of our lincats is used to 
implement syntactical functions, let us go through the 
various ways of constructing clauses:

\begin{lstlisting}
PredVP np vp = {s = \\t,a,p =>
  np.s!Nom
  ++ vp.subjpost 
  ++ verbStr vp t a p np.agr
  ++ vp.compl ;
} ;
\end{lstlisting}

The function §PredVP : NP -> VP -> Cl§ from module 
§Sentence§ is the basic way of constructing a clause from a 
subject and a verb phrase (23 out of the 31 overloads 
of §mkCl§ from the resource grammar API are implemented in terms 
of §PredVP§). In our implementation this function concatenates 
four elements:

\begin{enumerate}
  \item The nominative form of the subject noun phrase,
  \item any particles indicating the mood of the verb phrase 
  (such as ``nem'' indicating passive voice, or ``laz''/``eth'' corresponding 
  roughly to ``can''/``must'' in English)
  \item the verb inflected for tense, anteriority, 
  polarity and person/number of the noun phrase,
  \item and the complement of the verb phrase.
\end{enumerate}

Conjugation of the verb is implemented in the function 
§verbStr§ defined in the module §ResDot§:
\begin{lstlisting}
verbStr : Verb -> Tense -> Anteriority -> Polarity -> Agr -> Str = \v,t,a,p,agr -> let vf = (tapaToVForm t a p agr) in 
  case <t,a> of {
    <Present,Anter> => "ray" ;    -- the perfect tense marker
    _ => []   
  } ++
  case p of {
    Pos => []  ;
    Neg => "vos" 
  } ++
  v.s!vf ;
\end{lstlisting}

The argument §vp§ in §PredVP§ above is of type §VP§ which is 
defined as §VP = Verb ** {compl : Str; subjpost : Str}§. 
GF has a mostly structural (as opposed to nominative) 
type system and thus §VP§ is automatically subtype of §Verb§, 
and variables of §VP§ type can be used where a §Verb§ is expected
(as we do above).

The function §verbStr§ is reused in other functions concerned 
with the construction of clauses.

\begin{lstlisting}
SlashVP np vpsl = {
  s = \\t,a,p => verbStr vpsl t a p np.agr ;
  subj = np.s!Nom ++ vpsl.subjpost;
  objCase = vpsl.objCase
} ;
\end{lstlisting}

The function §SlashVP : NP -> VPSlash -> ClSlash§ is used 
in several overrides of the API functions §mkClSlash§ as well as 
in some overrides of §mkQCl§ and §mkRCl§. Since §VPSlash§ is 
also an extension of §Verb§, we can reuse §verbStr§ to 
conjugate the verb. The subject phrase is not prepended 
to the verb (for reasons explained in Section \ref{sec-verbs}), but instead 
kept in the field §subj§. Unlike §VP§, a §VPSlash§ has no complement, 
but the case of the missing object has to be remembered.

\begin{lstlisting}
RelVP rp vp = {
  s = \\t,a,p,anim,n =>
    rp.s!(anToQuForm anim n)!Nom
    ++ vp.subjpost
    ++ verbStr vp t a p (Ag P3 n)
    ++ vp.compl ;
} ;
\end{lstlisting}

The function §RelVP : RP -> VP -> RCl§ is the basic way of 
creating relative clauses in which the subject of the sentence 
is being relativized (22 of the 25 overrides of §mkRCl§ are 
based on §RelVP§). This implementation is quite similar 
to §PredVP§, but with the relative pronoun taking the place 
of the subject, and being inflected for animacy and number.

The verb agrees in number with the relative pronoun and 
is always in the third person.

Finally, clauses can be formed using §ExistNP : NP -> Cl§.

\begin{lstlisting}
ExistNP np = {s = \\t,a,p => 
  np.s!Gen
  ++ (verbStr vekhat_V t a p np.agr)
} ;
\end{lstlisting}

This is for creating sentences such as ``Lajaki vekha'' = ``There is a warrior''. 
Unlike any other verb in Dothraki, the verb ``vekhat'' assigns 
genitive case 
to the subject of the sentence. Apart from that, 
we again use §verbStr§ to conjugate the verb ``vekhat'' 
(§vekhat_V§ is defined in module §ExtraDot§).

\subsection{Vocabulary Extraction}\label{sec-vocab}

\subsubsection{Dictionaries and Lexicons in the RGL}
As mentioned earlier, many of the resource grammars in GF's
library also come with a dictionary often called \name{Dict\emph{XXX}}.

For example, in \name{DictGer} from the German resource grammar, we
find entries such as:
\begin{lstlisting}
entziehen_V = irregV "entziehen" "entzieht" "entzog" "entz`\"o`ge" 
  "entzogen" ; 
kreis_N = mkN  "Kreis" "Kreise" masculine ;
\end{lstlisting}
An application grammar writer can then use these definitions and does not
have to worry about the gender of nouns, plural forms 
or irregular conjugations.

In fact, the dictionaries in the RGL come in the
form of an abstract grammar with a concrete implementation. The
abstract grammar, often called \name{Dict\emph{XXX}Abs}, contains
entries like
\begin{lstlisting}
entziehen_V : V ;
kreis_N : N ;
\end{lstlisting}

These dictionaries, of course, cannot be used directly for translation.
However, there are two ways in which the RGL supports direct
translation between two langauges. First, there is the abstract
grammar \name{Lexicon} in the \name{abstract} folder, which covers
about 350 more or less commonly used words:
\begin{lstlisting}
...
drink_V2 : V2 ;
dry_A : A ;
dull_A : A ;
dust_N : N ;
...
\end{lstlisting}
This abstract grammar is then implemented in many of the languages
of the RGL. From \name{LexiconGer}:\footnote{The choice of names here is unfortunate:
In the common linguistic sense of the words, \name{DictGer} is a lexicon, while
\name{LexiconGer} is a dictionary...}
\begin{lstlisting}
drink_V2 = dirV2 Irreg.trinken_V ;
dry_A = regA "trocken" ;
dull_A = regA "stumpf" ;
dust_N = reg2N "Staub" "St`\"a`ube" masculine ;
\end{lstlisting}

The \name{Lexicon} module is part of the abstract grammar
\name{Lang} and its concrete implementations
in various languages. Of course, the vocabulary is far
too small to be useful for applications, but it is useful for
experimentation and for testing the resource grammars.

The RGL also has a folder \name{translator} which
mirrors this structure on a much larger scale:
The abstract grammar \name{Dictionary} contains 65,000 English 
words,\footnote{as well as 450 Finnish proper names...} and
the concrete implementations \name{Dictionary\emph{XXX}} contain
their translations into various languages. This is used, e.g.,
in the GF online translator \cite{gf-translate}.

\subsubsection{The Dothraki dictionary}

The publicly available vocabulary list for Dothraki \cite{vocab} currently
contains about 1,500 words.\footnote{This is still not enough to implement
the abstract \name{Lexicon} module from GF's resource grammar library:
Many common words do not yet have a Dothraki equivalent.}
In order to make these accessible for GF application grammars
and for the purpose of translation from and to English, we
supply, as part of our project, a Python script that translates
the wiki source of this list into a GF-compatible format.

The script creates five files in total, adhering closely 
to the structure established in the full RGL, as explained above:
\begin{itemize}
  \item the abstract grammar \name{DictDotAbs.gf} and 
  its concrete implementation \name{DictDot.gf} in the
  folder \name{simpledothraki},
  \item the abstract grammar \name{BigLexicon.gf} in
  \name{simpleabstract} and its implementations
  \name{BigLexiconEng.gf} and \name{BigLexiconDot.gf} in
  \name{simpleenglish} and \name{simpledothraki}, respectively.
\end{itemize}

\begin{figure}
\begin{tabular}{l|l}
\begin{minipage}{.44\linewidth}
\begin{lstlisting}
*'''aranikh''' [a?a'nix]
:''ni.'' subject, topic, theme

*'''gerat''' [ge'?at]
:''vtr.&rarr;abl.'' to lack
:{{Vsup|ger}}
\end{lstlisting}
\end{minipage}&
\begin{minipage}{.55\linewidth}
\begin{lstlisting}
`\textsf{in simpledothraki/DictDotAbs.gf}:`
aranikh_N : N ;
gerat_V2 : V2 ;

`\textsf{in simpledothraki/DictDot.gf}:`
aranikh_N = mkN "aranikh" inanimate ;
gerat_V2 = mkV2 "gerat" "ger" ablative ;

`\textsf{in simpleabstract/BigLexicon.gf}:`
subject_N : N ;
topic_N : N ;
theme_N : N ;
lack_V2 : V2 ;

`\textsf{in simpledothraki/BigLexiconDot.gf}:`
subject_N = DictSimpleEng.subject_N;
topic_N = DictSimpleEng.topic_N;
theme_N = DictSimpleEng.theme_N;
lack_V2 = DictDot.gerat_V2;

`\textsf{in simpleenglish/BigLexiconEng.gf}:`
subject_N = DictDot.aranikh_N;
topic_N = DictDot.aranikh_N;
theme_N = DictDot.aranikh_N;
lack_V2 = DictSimpleEng.lack_V2;
\end{lstlisting}
\end{minipage}
\end{tabular}
\caption{\label{fig-vocab}Two vocabulary wiki entries and generated GF definitions}
\end{figure}

Figure \ref{fig-vocab} shows an example of the GF code the script
generates. A word is added to \name{BigLexicon} whenever a
corresponding word is found in \name{DictEng}, the massive
English dictionary that is part of the RGL.

The wiki vocabulary list has 1,380 entries at the time
of this writing. From these, our script creates
1,410 definitions in \name{DictDot}.\footnote{the somewhat larger
number comes from the fact that many words can be used
in two different ways, for example both as an intransitive
and as a transitive verb, giving more than one entry in \name{DictDot}}

There are a total of 1,916 English translations in the vocabulary list.
The script is able to extract 1,059 entries for \name{BigLexicon}
from this. Many translations cannot be directly added to
\name{BigLexicon} for various reasons:
\begin{itemize}
  \item Some English words appear as translations for
  more than one Dothraki word. The script picks the first
  one it finds.\footnote{Obviously, that might not result in the
  best or most appropriate translation being picked, but that issue
  can be addressed by later manual editing of the generated list.}
  \item Many Dothraki words do not correspond to any one English word.
  For example the list contains entries for ``defensive sword strike'',
  ``sound of fire going out'', ``roast quail'' or
  ``time period of approximately two minutes''. While for some of these,
  it might be desirable to have dictionary 
  entries,\footnote{For example, there could be an entry
  §roast_quail_CN : CN§ in \name{BigLexicon} with linearizations
  §roast_quail_CN = mkCN DictEng.roast_A (mkCN DictEng.quail_N)§
  in
  \name{BigLexiconEng} and §roast_quail_CN = mkCN DictDot.mechikh_N§
  in \name{BigLexiconDot}.} this process is almost impossible
  to automate and must be done by hand.
\end{itemize}

While \name{DictDot} and \name{DictDotAbs} are meant to be used
in application grammars, the \name{BigLexicon\emph{XXX}} files
can be used for testing and translation. An example of this
is shown in the next section.

\subsection{Example}

With the current implementation status, it is already possible 
to parse and translate a wide range of phrases. For experimentation
and testing we supply the \name{LangSimple\emph{XXX}} abstract and
conrete grammars which simply import the \name{SyntaxSimple\emph{XXX}}
and \name{BigLexicon\emph{XXX}} modules (completely
analogous to the \name{Lang\emph{XXX}} module in the
full RGL) and can be used directly
from the GF command line:

\begin{lstlisting}
> i simpleenglish/LangSimpleEng.gf
linking ... OK

Languages: LangSimpleEng
2745 msec
LangSimple> i simpledothraki/LangSimpleDot.gf
linking ... OK

Languages: LangSimpleDot LangSimpleEng
187 msec
LangSimple> p "the warrior which protected the turtle has stabbed my goat with a dagger ." | l
lajak fin viljazer adraes ray vinde dorv anni ma mihesofoon .
the warrior which protected the turtle has stabbed my goat with a dagger .
lajak fin viljazer adraes ray vinde dorv ma mihesofoon anni .
the warrior which protected the turtle has stabbed my goat with a dagger .

31 msec
\end{lstlisting}

The multiple outputs for the last command are 
quite typical and derive from the fact that the parse of the English 
sentence is ambiguous:
Was the duck stabbed with a dagger or was a duck with a dagger stabbed?
Note that, in this case, the Dothraki translations differ, because the
possessive pronoun ``anni'' follows the noun phrase
(either ``dorv'' = ``goat'' or ``dorv ma mihesofoon'' = ``goat with a dagger'').

\section{Discussion}\label{sec-discuss}

To conclude our project documentation, we will now discuss problems 
and limitations as well as briefly describe future plans before 
we summarize the project in its current state and our experience 
while working on it.

\subsection{Difficulties}
Beginning to write a resource grammar for the first time was 
not easy. Due to the structure of the full RGL, it is not 
possible to implement a resource grammar for a new language
step-by-step, adding linearizations for a few important
categories and expanding gradually, at least not when basing
it directly on the abstract specifications from the full RGL.

Of course, it is not feasible either to implement the entire
resource grammar before being able to do any kind of testing.
As a remedy for this problem, Ranta recommends
taking the resource grammar of a similar language as a
starting point and changing it step 
by step \cite[p.\ 224]{gf-book}. 
This proved difficult in our case, since Dothraki, as a constructed
language, of course does not have family ties to any existing language. 
While its grammar is certainly inspired by phenomena seen in
natural languages, these appear to be picked from a variety
of different language families (many of which, as non-linguists, we are not
particularly familiar with). Furthermore, taking
a language like English as a starting point, which is certainly similar
to Dothraki in many aspects, but wildly different in others,
proved to be fraught with difficulties as well. This is because
of the web of dependencies in a grammar,
changes at one point, such as increasing the number of cases 
from two to five, tend to require changes in a number
of different places all over the grammar, an effect that
can quickly snowball. Finding our own solution for
this problem, namely growing the abstract part gradually 
(by selectively copying over parts from the full RGL) along with
the concrete implementation, took time, but in the end we 
are satisfied with how it turned out.

\subsection{Limitations}
As Ranta notes in \cite{gf-slides}, the most important
sources for creating a resource grammar are
``[a] \emph{good} grammar book'' and ``[a] \emph{good} dictionary''.

That poses some problems in the case of Dothraki.
There are basically only two sources of information
on Dothraki grammar: The writings of inventor 
David J.\ Peterson \cite{peterson-web},
the community-operated ``Tongues of Ice and Fire`` wiki
at \url{wiki.dothraki.org}. These tend to be great resources
for the basic parts of the language. Peterson's
blog contains detailed, accessible explanations of
many grammatical phenomena featured in
the snippets of Dothraki, that have appeared on the show
up to its third season. And the enthusiastic Dothraki community
has done great work organizing and detailing this information.

However, as our work progressed, it has become harder and harder
to find good information on some of the more subtle or
obscure parts of the grammar. It should be noted that the
total corpus of text in the Dothraki language 
still comprises only a few hundred sentences.
By now, we have already mostly
incorporated the information available from the two sources
mentioned above (the two largest remaining omissions are wh-questions and
numerals). In many cases we have run up against the limits
of what information is available and sometimes we have been
forced to resort to the Dothraki corpus itself
in an effort to try and reverse engineer
language features that so far have not been explained by
their inventor.

We give three examples:
\begin{itemize}
  \item As noted in Section \ref{sec-verbs} our resource grammar
  so far only contains a dummy implementation of
  verbs with a verb phrase complement (§VV§s). In English, 
  (auxiliary) verbs like ``can'', ``must'' and ``want to'' are
  of this kind. In Dothraki, as explained in the quoted section,
  ``can'' and ``must'' are rendered as particles
  (``laz'' and ``eth'') and this is what our current
  implementation captures. However, certainly not all
  §VV§s are rendered like this
  in Dothraki. These non-inflected particles almost surely
  are a small closed class of words, whereas §VV§s are
  plentiful (think ``agree to (doing sth)'', ``consider (doing sth)'',
  ``enjoy (doing sth)'', ``expect to (do sth)'' etc.).
  However, there is almost no explanation available
  on the syntactical structure of such verbs in Dothraki.

  When they have appeared on the show, verb phrase
  complements are usually avoided in favour of
  embedded clauses:

  \begin{quote}
  ``The khaleesi wants to eat something different tonight.'' =
  ``Khaleesi zala meme adakha esinakh ajjalan.'' (lit.\ ``The khaleesi wants that she eats something different tonight'')\\
  ``You'd rather be sold into slavery?'' = ``Yer zali meyer nem vazhi ven zafra?'' (lit.\ ``You want that you are sold into slavery?'')
  \end{quote}

  As far as we can tell, in the entire corpus there are only four sentences
  in which the verb ``zalat'' (to want) is used with what seems to be a
  true verb phrase complement:
  \begin{quote}
    ``Hash shafka zali addrivat mae [...]?'' = ``Do you want me to kill him?''\\
    ``[...] hash yer zali zifichelat moon?'' = ``[...] you want to steal from him?''\\
    ``Anha zalak nesat ven fini ven athyazhar khaleesisi.'' = ``I'd like to know what a Khaleesi tastes like.''\\
    ``Hash yeri vo zali nesat rek dirgak anha?'' = ``Don't you want to know what I think?''
  \end{quote}
  We could not find any other instances of §VV§ verbs being used
  with a true verb phrase complement in the corpus.

  From these four sentences it certainly appears like verb phrase complements
  (similar to English) are formed with the verb in the infinitive
  (``addrivat'', ``zifichelat'' and ``nesat'' are all infinitive forms).
  However, as far as we can tell, this has not been discussed by 
  Peterson so far, nor is there any information about it on the wiki.

  It seems rather premature and speculative
  to conclude from just four sentences (all featuring
  the same verb) that this is the general way verb phrase
  complements are rendered in Dothraki. Therefore, we are at an
  impasse: Until further information becomes available,
  we cannot implement §VV§s (other than the few
  special ones like ``can'' and ``must'').

  \item The module \name{Structural} contains a number
  of common structural words that should be available in 
  every language. Quite a few of these are simply not yet
  available in Dothraki (``somewhere'', ``almost'',
  ``quite'', ``otherwise'', ``at least'' etc.), but
  even those that have translations in the vocabulary list
  are sometimes difficult to implement. For example, the
  vocabulary list notes that ``someone'' translates as
  ``ato''. But since this is a noun phrase, it probably
  must inflect somehow.
  Inflection in nouns depends on their animacy, but
  it is not clear whether inanimate or animate declension
  is used here (or some irregular inflection, as with determiners). 
  In the corpus ``ato'' appears exactly once,
  in the nominative. In the end, we had no other choice
  but to guess the declension (we decided on standard
  animate inflection).

  Similarly, ``eyak'' supposedly means ``everybody'', according
  to the vocabulary, but in the only place it is used
  in the corpus, it seems to be used in the sense of ``everything''.
  In that instance it is used in the genitive (probably; the phrase
  is ``hatif eyaki'' and the preposition ``hatif'' can assign
  genitive, allative or ablative case, depending on meaning).
  In the end we decided on a declension based on this one occurence and
  by analogy with ``vosak'' (= ``nobody''), which occurs in
  the accusative once.

  In the end, in many cases in the \name{Structural} module, 
  we had to employ something more akin to educated guesswork rather
  than being able to rely on definitive sources.

  \item §Comp§s (``complementizers to a copula'') can be used
  to form imperatives: ``Be strong!'' or ``Be a warrior!''.
  The first has an obvious rendering in Dothraki (by putting
  the verb ``hajat'' = ``to be strong'' derived from ``haj'' = ``strong''
  in the imperative), but the second does not. In Dothraki
  there is no analogue of the copula ``to be'' that could be
  used in the imperative here. In the corpus, no sentence
  like this exists. For now, we cannot implement this
  correctly.
\end{itemize}

We are hopeful that, in spite of these current limitations,
our work so far can serve as a good basis for implementing
more language features as further and more precise information 
on them becomes available.

\subsection{Outlook}
In the future, we would like to expand our resource grammar to a
full resource grammar and add it to the GF resource grammar library.
We are already working on and hosting the project online, for everyone
to see, so that the GF and/or Dothraki community may help us to achieve
this goal. Some more parts we can implement right now, for some we have
to wait until more information regarding the Dothraki language is being
published. In the end, we also want to add the full vocabulary to our
dictionary.

\subsection{Conclusion}
Now that the first milestone of our project lies behind us,
we can evaluate the working process and the result. As we had
some starting problems at the beginning of the project, we are
proud to have been able to create such a big portion of the
resource grammar in the given time. By relying on the basics
learned during the seminar and by aiming for such an ambitious
goal, we surely learned a lot during the process. As we are
convinced that our project can be helpful and interesting for
other people in the future, we were motivated to keep on working
on the resource grammar and to try to find the best solution for
every problem. This term paper does not mark the end of this project,
but we are curious to hear what you, our lecturers, and other people
think of the work we have done so far.

\backmatter

\section*{Appendix}

\paragraph{Source Code}

The complete source code can be viewed and downloaded via GitHub: 
\url{https://github.com/mahen2/dothraki_gf}

\makeatletter
\renewcommand\@biblabel[1]{}
\renewenvironment{thebibliography}[1]
     {\section*{\refname}%
      \@mkboth{\MakeUppercase\refname}{\MakeUppercase\refname}%
      \list{}%
           {\leftmargin0pt
            \@openbib@code
            \usecounter{enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}
\makeatother

\begin{thebibliography}{9}
\bibitem[Brown \& Frederking, 1995]{brown-frederking}
Brown, R.\ \& Frederking, R.\ (1995). 
\emph{Applying statistical English language modeling to symbolic machine 
translation}. 
In Proceedings of the 
sixth international conference on theoretical and methodological 
issues in machine translation (TMI-95), pp. 221--239.

\bibitem[Cognitus Apps, 2016]{cognitus}
Cognitus Apps (2016). 
Flamingo Dothraki. 
Retrieved from \url{https://play.google.com/store/apps/details?id=com.cognitusapps.flamingo.dothraki&hl=de}

\bibitem[Fun Translations, 2016]{funtrans}
Fun Translations (2016). Dothraki translator. 
Retrieved from \texttt{http://funtranslations.com/dothraki}

\bibitem[D\'etrez\ \& Camilleri, 2016]{detrez-camilleri}
D\'etrez, G.\ \&\ Camilleri, J.J.\ (2016). \emph{Grammatical Framework}. 
GitHub.com. Retrieved from \texttt{https://github.com/GrammaticalFramework}

\bibitem[Google, 2016]{google}
Google (2016). \emph{Google Translate}. Retrieved from 
\url{https://translate.google.com/}

\bibitem[Google Groups, 2016]{google-groups}
Google Groups (2016). \emph{Problem with strange metavariables in parsing}.
Retrieved from \url{https://groups.google.com/d/msg/gf-dev/aRjt_2JfvA0/iTcZbTu5AAAJ)}

\bibitem[Grammatical Framework, 2016a]{gf-translate}
Grammatical Framework (2016a). 
\emph{Translation with GF: Powered by multilingual grammars}. 
Retrieved from \url{http://www.grammaticalframework.org/demos/translation.html}

\bibitem[Grammatical Framework, 2016b]{gf-rgl}
Grammatical Framework (2016b). 
\emph{GF Resource Grammar Library: Synopsis}. 
Retrieved from \url{http://www.grammaticalframework.org/lib/doc/synopsis.html}

\bibitem[Martin, 2016]{martin}
Martin, G.R.R. (2016). \emph{Bibliography}. 
Retrieved from \url{http://www.georgerrmartin.com/bibliography/}

\bibitem[Henkel \& Kuckuck, 2016]{github}
Henkel, M.\ \& Kuckuck, B.\ (2016). 
\emph{Building a resource grammar for Dothraki}. GitHub.com. 
Retrieved from \url{https://github.com/mahen2/dothraki_gf}

\bibitem[Peterson, 2014]{living-language}
Peterson, D.J.\ (2014). 
\emph{Living Language Dothraki: A Conversational Language Course 
Based on the Hit Original HBO Series Game of Thrones.}
New York: Living Language.

\bibitem[Peterson, 2015]{peterson-book}
Peterson, D.J.\ (2015). \emph{The Art of Language Invention.}
Penguin Books.

\bibitem[Peterson, 2016]{peterson-web}
Peterson, D.J.\ (2016). \emph{About Dothraki}. 
Retrieved from \url{http://www.dothraki.com/about-dothraki/}

\bibitem[Ranta, 2009]{ranta-rgl}
Ranta, A.\ (2009). 
\emph{The GF Resource Grammar Library.}
Linguistic Issues in Language Technology, 2(2), pp.\ 1--63.

\bibitem[Ranta, 2011a]{gf-book}
Ranta, A.\ (2011a). 
\emph{Grammatical Framework: Programming with Multilingual Grammars.}
Stanford: CSLI Publications.

\bibitem[Ranta, 2011b]{gf-slides}
Ranta, A.\ (2011b). 
\emph{Grammatical Framework: Programming with Multilingual Grammars. 
Slides for the GF book}. Retrieved from 
\url{http://www.grammaticalframework.org/gf-book/gf-book-slides.pdf}

\bibitem[Ranta, 2015]{gf-status}
Ranta, A.\ (2015). 
\emph{The Status of the GF Resource Grammar Library}. Grammatical Framework. 
Retrieved from \url{http://www.grammaticalframework.org/lib/doc/status.html}

\bibitem[Tongues of Ice and Fire Wiki, 2016]{vocab}
Tongues of Ice and Fire Wiki (2016). \emph{Vocabulary}. 
Retrieved from \url{http://wiki.dothraki.org/Vocabulary}

\bibitem[Wikipedia, 2016]{wiki-gf}
Wikipedia (2016). \emph{Grammatical Framework}.
Retrieved from \url{https://en.wikipedia.org/wiki/Grammatical_Framework}
\end{thebibliography}

\end{document}